Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-26 paper_jasa.tex
p-value from definition
subject specific power corve updated

\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data. \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for p-value.  \\ Should I change everything accordingly? I see that we have been careful not to use $\beta$ for power! }

@@

{\bf Screening scenarios:} \hh{Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.  In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.} 
%There are six suggested approach to clean the data. 
They are 
\begin{enumerate}
\item {\bf include all participants} and their evaluations
\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records. If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
\end{enumerate}

@@

@@

Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2. For experiment 3 the scenario is different since visually we can easily notice this type of contamination in the data. Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$. While fitting model \ref{mixed} we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated. Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2. For experiment 3 we fit the intercept two allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.

@@

\green{third experiment doesn't really have a p-value, does it? Need to explain this. Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.}

@@

Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}. From Equation \ref{fig:pval_plot_signal}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $. This gives the signal strength estimate as  
\[
\hat p_D= (1- {u}/{K})/({m-1})
\]
where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations. 
Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values. 
\hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$. This estimate has to be interpreted carefully, since the  choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense. Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.  }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}. But it is true for experiment 2, right?. }
%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.

@@

>>>>

\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate observed data. \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.  \\ Should I change everything accordingly? I see that we have been careful not to use $\beta$ for power! }

@@

{\bf Screening scenarios:} Amazon Mechanical Turk data often suffers from the problem that some participants are trying to 'game' the system -- in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.  In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis. 
%There are six suggested approach to clean the data. 
They are
\begin{enumerate}
\item {\bf include all participants} and their evaluations
\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records. If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
\end{enumerate}

@@

Table \ref{tbl:summary} shows  the summary results after applying different screening on the data. For our analysis we picked criteria six (6) and present results based on this in the following sections.

@@

Figure \ref{fig:power_loess_effect} also reveals that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiment 1 and 2. For experiment 3 the scenario is different since contamination in the data is visible and this sets the lower limit of proportion correct at a higher value. Thus for fitting model \ref{mixed} to the data we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$. While fitting model \ref{mixed} we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated. Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2. For experiment 3 we fit the intercept to allow the power curves to go above 0.05 as we see in figure \ref{fig:power_loess_effect}.

@@

\green{third experiment doesn't really have a p-value, does it? Need to explain this. Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway. So many small p-values are coming from the fact that contamination makes the slope negatively significant. For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant. It is like the effect of influential outlier.}

@@

Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}. From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $. This gives the signal strength estimate as  
\[
\hat p_D= (1- {u}/{K})/({m-1})
\]
where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations. 
Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values. 
\hh{Note that the estimate for $p_D$ is bounded within $(0, 1/(m-1))$. This estimate has to be interpreted carefully, since the  choice of a large lineup size, $m$, will make the value look `significant' in a traditional sense. Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.}
\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.48, 0.85, and 0.11 based on $n_1 = 39, n_2 = 41$, and $n_3= 18$ lineups, respectively.  }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}. But it is true for experiment 2, right?. }
%is equivalent to classical test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . p-values are shown on square root scale scale. }
    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
       \label{fig:pval_definition}
\end{figure*}
